---
title: "Explore Mixtures"
output: 
  bookdown::word_document2:
    fig_caption: yes
    toc: yes
editor_options: 
  chunk_output_type: console
---

Here we explore land use and chemical mixture relationships.



```{r setup, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.width = 7,
                      fig.height = 5)

library(tidyverse, quiet = TRUE)
library(rpart, quiet = TRUE)
library(party, quiet = TRUE)
library(partykit, quiet = TRUE)
library(rpartScore, quiet = TRUE)
library(readxl, quiet = TRUE)
library(flextable, quiet = TRUE)
# library(subselect) #sim annealing
library(leaps, quiet = TRUE) #regsubsets
library(glmnet, quiet = TRUE)
library(toxEval)

source("R/mixtures/mix_script.R")
drake::loadd(chemicalSummary)
drake::loadd(chem_info)
```

```{r getData}
panther <- data.table::fread("panther_data/joined_genes.csv")
end_point_info <- end_point_info

panther_df <- panther %>% 
  select(gene = gene_abbr,
         pathway_accession,
         pathway_name) %>% 
  left_join(select(end_point_info,
                   endPoint = assay_component_endpoint_name,
                   gene = intended_target_gene_symbol),
            by="gene")

chemicalSummary_panther <- chemicalSummary %>% 
  left_join(panther_df, by="endPoint") %>% 
  select(-endPoint) %>% 
  rename(endPoint = pathway_name) %>% 
  filter(!is.na(endPoint),
         endPoint != "")

ear_thresh <- 0.001
n_site_thresh <- 4

```

We will look at mixtures based on unique chemical/endpoint combinations. The way we prioritize mixutres is summing EARs for each endpoint, and then only look at endpoints that have at least an sumEAR of `r ear_thresh`. Then, we go back and figure out which mixtures occur at at least `r n_site_thresh` sites. The only mixtures that fit those 2 critiera are:

There are `r length(unique(filter(chemicalSummary, EAR>0) %>% pull(CAS)))` detected chemicals associated with ToxCast endpoints. There are `r length(unique(filter(chemicalSummary_panther, EAR>0) %>% pull(CAS)))` detected chemicals associated with Panther pathways. The chemicals that we lose are:

```{r missingChems, results='asis'}
missing_chemicals <- chemicalSummary %>% 
  filter(EAR > 0) %>% 
  filter(!(CAS %in% unique(chemicalSummary_panther$CAS))) 

plot_tox_boxplots(missing_chemicals, category = "Chemical")

chnm_miss <- missing_chemicals %>% 
  select(chnm) %>% 
  mutate(chnm = as.character(chnm)) %>% 
  distinct() %>% 
  pull(chnm)

# chnm_miss

```

```{r getMixtures, results="asis"}

cs <- chemicalSummary_panther


EARsum_endpoint <- sum_endpoints(cs, ear_cutoff = ear_thresh)

contributing_chems <- calc_contr_chems(EARsum_endpoint)

chm_key <- contributing_chems %>% 
    select(contr_chems) %>% 
    distinct() %>% 
    left_join(unique(select(contributing_chems,
                            contr_chems,
                            contr_chems_lt)), 
              by="contr_chems")

top_mixes <- contributing_chems %>% 
    group_by(contr_chems, endPoint) %>% 
    summarise(n_samples = n(),
              unique_sites = length(unique(site))) %>% 
    filter(unique_sites > {{n_site_thresh}}) %>% 
    arrange(desc(n_samples)) %>% 
    left_join(chm_key, by="contr_chems") %>% 
    mutate(contr_chems_st =
             paste(unique(unlist(contr_chems_lt)),
                   collapse = ",")) %>% 
  filter(!duplicated(contr_chems))

# Let's only take the top endPoint of single chem mixtures:


mix_graph_data <- data.frame()

for(i in 1:nrow(top_mixes)){
    
    chem <- top_mixes$contr_chems_lt[i]
    endpoint <- top_mixes$endPoint[i]
    
    mix_EAR_sum <- cs %>%
        mutate(chnm = as.character(chnm)) %>% 
        filter(chnm %in% unlist(chem),
               endPoint == endpoint) %>% 
        group_by(site, date, endPoint, shortName) %>% 
        summarise(sumEAR = sum(EAR)) %>%
        mutate(mix_st = paste(unlist(chem),
                              collapse = ","),
               mix = list(unlist(chem)),
               mixID = i)
    
    if(length(unlist(chem)) == 1){
      mix_EAR_sum$mix_st <- paste0(mix_EAR_sum$mix_st,
                                   " (", mix_EAR_sum$mixID, ")")
    }
    
    mix_graph_data <- bind_rows(mix_graph_data,
                                mix_EAR_sum)
}

priority_endpoints <- top_mixes$endPoint

mixtures_in_ep <-
  cs %>% 
  filter(endPoint %in% priority_endpoints) %>% 
  group_by(site, shortName, date, endPoint) %>%
  mutate(sum_ear_endpoint = sum(EAR)) %>%
  ungroup() %>%
  mutate(chem_mix_contribution = (EAR/sum_ear_endpoint)*100) %>% 
  filter(chem_mix_contribution > 1) %>% 
  group_by(site, shortName, date, endPoint, sum_ear_endpoint) %>%
  summarize(n_contr_chems = n(),
            contr_chems_lt = list(as.character(unique(chnm))),
            contr_cas_lt = list(as.character(unique(CAS))),
            contr_chems_st = paste(as.character(unique(chnm)),
                                collapse = ","),
            contr_chems = paste(as.character(unique(CAS)),
                                collapse = ","),
            max_individual_contr = max(EAR)) 

key <- mixtures_in_ep %>% 
              ungroup() %>%
              select(contr_chems,
                     contr_chems_st,
                     contr_chems_lt) 

key <- key[!duplicated(key$contr_chems), ]

top_mixes_per_ep <- mixtures_in_ep %>% 
  group_by(endPoint, contr_chems) %>% 
  summarize(n_sites = length(unique(site)),
            sumEAR = sum(sum_ear_endpoint)) %>% 
  filter(sumEAR == max(sumEAR)) %>% 
  ungroup() %>% 
  left_join(key,
            by="contr_chems")

# Mixtures to look at:
check_mix <- top_mixes %>% 
  ungroup() %>% 
  select(chemicals = contr_chems_st,
         chem_list = contr_chems_lt) %>% 
  bind_rows(select(top_mixes_per_ep, 
                   chemicals = contr_chems_st,
                   chem_list = contr_chems_lt))

if(all(c("Atrazine,DEET", "DEET,Atrazine") %in% check_mix$chemicals)){
  check_mix <- check_mix[check_mix$chemicals !=
                         "Atrazine,DEET", ]  
}

if(all(c("Benz(a)anthracene,Indeno(1,2,3-cd)pyrene",
         "Indeno(1,2,3-cd)pyrene,Benz(a)anthracene") %in% check_mix$chemicals)){
  
  check_mix <- check_mix[check_mix$chemicals !=                        "Indeno(1,2,3-cd)pyrene,Benz(a)anthracene", ]  
}

check_mix <- check_mix[!duplicated(check_mix$chemicals),]

top_mixes_select <- top_mixes %>% 
  ungroup() %>% 
  select(chemicals = contr_chems_st, 
         endPoint, n_samples, unique_sites) %>% 
  filter(!duplicated(chemicals))


mix_ft <- flextable(top_mixes_select)
autofit(mix_ft)

```

# Overview

```{r regularBP, fig.cap="Urban hits"}

df_lu <- open_land_use() %>% 
  filter(site %in% unique(cs$site))

mix_graph_data_lu <- mix_graph_data %>% 
    left_join(df_lu[,c("site","Urban","Agriculture")],
              by = "site") %>% 
    mutate(thresh = sumEAR > 0.001,
           facet = paste(mix_st, endPoint, sep = "\n"),
           mixture_id = as.integer(as.factor(facet)))

# TODO: actually us MDL EARS!!!

mix_graph_data_lu <- create_censor(mix_graph_data_lu)

n_fun <- function(x){
  return(data.frame(y = 0,
                    label = length(x)))
}

urban <- ggplot(data = mix_graph_data_lu,
                aes(x = thresh, y = Urban)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(color = "grey50", alpha = 0.4) +
    stat_summary(fun.data = n_fun,
                 geom = "text", hjust = 0.5) +
    facet_wrap(. ~ facet, ncol = 4) +
    theme_bw() +
    theme(strip.text = element_text(size = 8)) +
    xlab(paste0("EAR > ", ear_thresh))

# As you can see by Figure \@ref(fig:regularBP)
```




# Trees by Urban/Ag

```{r party2, results="asis"}

library(officer)
big_border = fp_border(color="orange", width = 2)
border_v = fp_border(color="gray")
border_h = fp_border(color="gray")

form_lm <- formula(sumEAR ~ Urban + Agriculture + Urban^2 + Agriculture^2)

form2_lm <- reformulate(termlabels = c("Urban","Agriculture"),
            response = "log10(highEAR)")

form_survival <- formula(survival::Surv(lowEAR,
                                highEAR,
                                 type="interval2") ~
                                 Urban + Agriculture +
                                 Urban^2 + Agriculture^2)

#The log goes into the distribution term in the model function:
form_survival_log <- formula(survival::Surv(lowEAR, 
                                highEAR,
                                 type="interval2") ~
                                 Urban + Agriculture +
                                 Urban^2 + Agriculture^2)

df_1 <- data.frame()

for(i in seq_len(nrow(top_mixes))){
    
  chem <- top_mixes$contr_chems_lt[i]
  endpoint <- top_mixes$endPoint[i]
  
  sub_df <- 
    mix_graph_data_lu %>% 
    filter(mixID == i) %>% 
    select(-mix)
    
  cat('\n##', unique(sub_df$mix_st), '\n')
  
  cat("\n### Tree\n")
  tree_return <- plot_trees(form_lm, sub_df, endpoint)

  cat("\n")
  
  cat("\n### Linear\n")
  x_df <- plot_lm(form_lm, form_survival, sub_df)
  
  cat("\n")
  cat("\n### Log\n")
  x_df2 <- plot_lm(form2_lm,
                   form_survival_log, 
                   sub_df, 
                   log=TRUE)
  cat("\n")
  cat("\n### Summarize variables\n")
  
  df_1 <- variable_summary(chem, endpoint, x_df, x_df2)
  table_mix <- flextable(df_1)
  table_mix <- width(table_mix, width = 1.5)
  table_mix <- border_remove(x = table_mix)
  table_mix <- border_outer(table_mix, 
                            part="all", 
                            border = big_border)
  table_mix <- border_inner_h(table_mix, part="all",
                            border = border_h )
  table_mix <- border_inner_v(table_mix, part="all",
                            border = border_v )
  cat(knitr::knit_print(table_mix))
}



cat("\n")

```

# Variable trimming

Variables available:

```{r table_of_categories}

df <- data.frame(a = names(df_lu)[2:12],
                 b = names(df_lu)[13:23],
                 c = names(df_lu)[24:34],
                 stringsAsFactors = FALSE)

autofit(flextable(df))

```

Variables that have at least 1 site with at least 2 percent:

```{r corLU, results="asis"}
library(corrplot)

big_enough <- function(x, thresh = 2){
  max(x, na.rm = TRUE) > thresh
}

df_lu_filtered <- df_lu %>% 
  select(-site) %>% 
  select_if(big_enough) 

df_lu_filtered <- df_lu_filtered%>% 
  bind_cols(select(df_lu, frac_2010, frac_2014))

M <- cor(df_lu_filtered)
corrplot(M, type="upper", tl.cex = 0.5)
```

Reduce the variables by those that have a corrolation coefficient > 0.9 or < -0.9.

```{r corLU_reduced, results="asis"}
lu_vars <- colnames(M)
lu_vars_rev <- rev(lu_vars)

exclude <- c()
exclude_rev <- c()

for(i in seq_along(lu_vars)){
  
  check_var <- lu_vars[i]
  
  if(check_var %in% exclude){
    next
  } else {
    check_cor <- names(which(M[,check_var] > 0.9 | 
                        M[,check_var] < -0.9))
    check_cor <- check_cor[check_cor != check_var]
    exclude <- c(exclude, check_cor)
  }

}

# for(i in seq_along(lu_vars_rev)){
#   
#   check_var_rev <- lu_vars_rev[i]
#   
#   if(check_var_rev %in% exclude_rev){
#     next
#   } else {
#     check_cor <- names(which(M[,check_var_rev] > 0.95 | 
#                         M[,check_var_rev] < -0.95))
#     check_cor <- check_cor[check_cor != check_var_rev]
#     exclude_rev <- c(exclude_rev, check_cor)
#   }
# 
# }

df_lu_filtered <- df_lu_filtered[,-which(names(df_lu_filtered) %in% exclude)]

M2 <- cor(df_lu_filtered)
corrplot(M2, type="upper", tl.cex = 0.5)
```

The variables left by the 0.9 threshold:

```{r resultingVars, results="asis"}
df <- data.frame(a = names(df_lu_filtered)[1:6],
                 b = names(df_lu_filtered)[7:12],
                 c = c(names(df_lu_filtered)[13:16],"",""),
                 stringsAsFactors = FALSE)
cat("\n")
ft <- flextable(df)
ft <- delete_part(x = ft, part = "header")
autofit(ft)
cat("\n")
```

Now, we use the sites with detections and filter out variables that don't have a single site left that doesn't have 10% of that category.

```{r corLU_manual, results="asis"}

df_lu_filtered <- bind_cols(df_lu[,c("site","Urban","Crops")],
                            df_lu_filtered) %>% 
  select(-Basin_Area_mi2, -Population_Density)

key2 <- cs %>% 
  select(chnm, Class) %>% 
  distinct() %>% 
  mutate(chnm = as.character(chnm),
         Class = as.character(Class))

check_mix$title <- NA

for(i in seq_len(nrow(check_mix))){
  chemicals <- unlist(check_mix$chem_list[i])
  class <- key2$Class[key2$chnm %in% unlist(chemicals)]
  check_mix$title[i] <- paste0(chemicals,":",class, collapse = "\n")
}

x <- data.frame(matrix(NA,
                  nrow = length(names(df_lu_filtered)), 
                  ncol = nrow(check_mix)))
names(x) <- check_mix$title

for(i in 1:ncol(x)){
  x[[i]] <- names(df_lu_filtered)
}
 
x <- x[-1,]

big_enough5 <- function(x, thresh = 10){
  max(x, na.rm = TRUE) > thresh
}

top_mixes$variables_to_use <- list(NA)
top_mixes_per_ep$variables_to_use <- list(NA)
# Get rid of variables with small influence:
for(i in 1:ncol(x)){
  
  chems <- unlist(check_mix$chem_list[i])
  
  sub_df <- cs %>% 
    mutate(chnm = as.character(chnm)) %>% 
    filter(chnm %in% chems) %>%
    filter(EAR > 0) 
  
  sites <- unique(sub_df$site)
  
  sub_df_lu <- df_lu_filtered %>% 
    filter(site %in% sites) %>% 
    select(-site) %>% 
    select_if(big_enough5)
  
  x[!(x[[i]] %in% names(sub_df_lu)),i] <- ""
  
  index_tm <- sapply(top_mixes$contr_chems_lt, function(x){
    all(x %in% chems) 
  })
  
  index_tm_ep <- sapply(top_mixes_per_ep$contr_chems_lt, function(x){
    all(x %in% chems) 
  })
  
  top_mixes$variables_to_use[index_tm] <- list(x[[i]][x[[i]] != ""])
  top_mixes_per_ep$variables_to_use[index_tm_ep] <- list(x[[i]][x[[i]] != ""])
  
}

x <- x[rowSums(x == "") != ncol(x), ]

# data.table::fwrite(x, "table_of_landuse.csv")

```


Now we can think about individual categories to remove. For now...I'm taking out Basin Area, pulling Crops back in, and substituting Urban for Population Density.



# Trees by more complicated land use options



```{r moreLandUse, results="asis"}

not_all_na <- function(x) all(!is.na(x))

mix_graph_data_lu <- 
  mix_graph_data_lu %>%
  select(-mix) %>% 
  left_join(select(df_lu_filtered,
                   -Urban, -Agriculture),
            by="site")

for(i in seq_len(nrow(top_mixes))){
    
  chems <- unlist(top_mixes$contr_chems_lt[i])
  mixture <- top_mixes$contr_chems_st[i]
  endpoint <- top_mixes$endPoint[i]
  
  variables_to_use <- top_mixes$variables_to_use[[i]]

  sub_df <- 
      mix_graph_data_lu %>% 
      filter(mixID == i) %>%
      distinct()
  
  cat('\n##', unique(sub_df$mix_st), '\n')
  
  cat("\n### Tree\n")
  form_bigger_lm <- formula(paste("sumEAR ~ ",
                               paste(variables_to_use,
                                     collapse = " + ")))
  
  tree_return <- plot_trees(form_bigger_lm, sub_df, endpoint)
  cat("\n")

  new_form_lm <- get_formula(sub_df, variables_to_use, 
                          lasso = FALSE, survival = FALSE)
  
  new_form_surv <- get_formula(sub_df, variables_to_use, 
                        lasso = FALSE)
  
  cat("\n### Linear\n")

  form_surv <- reformulate(termlabels = attr(terms(new_form_surv),"term.labels"),
              response = 'survival::Surv(lowEAR,
                                      highEAR,
                                      type="interval2")')

 x_df <- plot_lm(new_form_lm, 
                form_surv, 
                sub_df = sub_df)
 
 cat("\n")   

  cat("\n### Log\n")
  new_form2 <- get_formula(sub_df,
                           variables_to_use,
                           survival = FALSE,
                           sumEAR = "highEAR",
                           log=TRUE)
  new_form_surv <- get_formula(sub_df,
                         variables_to_use,
                         sumEAR = "highEAR",
                         log=TRUE)
  
  new_form_surv_log <- reformulate(termlabels = attr(terms(new_form_surv),"term.labels"),
              response = 'survival::Surv(lowEAR,
                                      highEAR,
                                      type="interval2")')

  x_df2 <- plot_lm(new_form2,
                   new_form_surv_log, 
                   sub_df,
                  log=TRUE)
  
  df_temp <- variable_summary(chems, endpoint, x_df, x_df2)
    
  cat("\n")
  cat("\n### Summarize variables\n")
  cat("\n")
  table_mix <- flextable(df_temp)
  table_mix <- width(table_mix, width = 1.5)
  table_mix <- border_remove(x = table_mix)
  table_mix <- border_outer(table_mix, 
                            part="all", 
                            border = big_border)
  table_mix <- border_inner_h(table_mix, part="all",
                            border = border_h )
  table_mix <- border_inner_v(table_mix, part="all",
                            border = border_v )
  cat(knitr::knit_print(table_mix))
  cat("\n")

}

```


# All EARs for our top endpoints:

```{r allEARs, results ='asis', eval=FALSE}

for(i in seq_len(nrow(top_mixes_per_ep))){
  
  ep <- top_mixes_per_ep$endPoint[i]
  chems <- top_mixes_per_ep$contr_chems_lt[[i]]
  
  cat('\n##', ep, '\n')

  cat('\n###', paste(chems, collapse = ", "), '\n')
  
  sub_df <- 
    cs %>% 
    mutate(chnm = as.character(chnm)) %>% 
    filter(endPoint %in% ep,
           chnm %in% chems) %>% 
    group_by(site, date, shortName) %>% 
    summarise(sumEAR = sum(EAR),
              mix_st = paste(unique(chnm),
                             collapse = ",")) %>%
    ungroup() %>% 
    left_join(df_lu_filtered, by="site") %>% 
    select_if(not_all_na)
  
  if(length(unique(sub_df$site)) < 10){
    break
  }
  
  sub_df <- create_censor(sub_df)
  
  variables_to_use <- top_mixes_per_ep$variables_to_use[[i]]

  form_bigger <- formula(paste("sumEAR ~ ",
                           paste(variables_to_use,
                                 collapse = " + ")))
  
  cat("\n#### Tree\n")
  tree_return <- plot_trees(form_bigger, sub_df, ep)
  cat("\n")

  cat("\n#### Linear\n")
  cat("\n")
  
  new_form_lm <- get_formula(sub_df,
                             variables_to_use,
                             survival = FALSE)
  new_form_surv <- get_formula(sub_df, variables_to_use)
  
  if(is.null(new_form_lm)){
    message("No linear")
    x_df <- data.frame()
  } else {
    new_form_surv <- reformulate(termlabels = attr(terms(new_form_surv), "term.labels"), 
    response = 'survival::Surv(lowEAR,
                               highEAR,
                               type="interval2")')
    x_df <- plot_lm(new_form_lm,
                    new_form_surv, sub_df)
  }
  
  cat("\n")
  cat("\n#### Log\n")
  cat("\n")

  new_form_lm_log <- get_formula(sub_df,
                              variables_to_use,
                              survival = FALSE,
                              sumEAR = "highEAR",
                              log = TRUE)
  new_form_surv_log <- get_formula(sub_df,
                            variables_to_use,
                            sumEAR = "highEAR",
                            log = TRUE)
  if(is.null(new_form_lm_log)){
    message("No log")
    x_df2 <- data.frame()
  } else {
    
    new_form_log_surv <- reformulate(termlabels = attr(terms(new_form_surv_log), "term.labels"), 
    response = 'survival::Surv(lowEAR,
                               highEAR, 
                               type="interval2")')
    
    
    x_df2 <- plot_lm(new_form_lm_log,
                     new_form_log_surv, sub_df,
                     log = TRUE)
  }

  cat("\n")
  
  df_temp <- variable_summary(chems, ep,
                              x_df, x_df2)
  
  cat("\n#### Summarize variables\n")
  cat("\n")
  table_mix <- flextable(df_temp)
  table_mix <- width(table_mix, width = 1.5)
  table_mix <- border_remove(x = table_mix)
  table_mix <- border_outer(table_mix, 
                            part="all", 
                            border = big_border)
  table_mix <- border_inner_h(table_mix, part="all",
                            border = border_h )
  table_mix <- border_inner_v(table_mix, part="all",
                            border = border_v )
  cat(knitr::knit_print(table_mix))
  cat("\n")

  
}


```

# Class Exploration

```{r classStuff}
class_sum <- cs %>% 
  left_join(df_lu_filtered, by = "site")

graph_class <- class_sum %>% 
  group_by(Class, site, date, Urban, Agriculture) %>% 
  summarize(sumEAR = sum(EAR, na.rm = TRUE)) %>% 
  group_by(Class, site, Urban, Agriculture) %>% 
  summarise(maxSumEAR = max(sumEAR)) %>% 
  ungroup()

ggplot(data = graph_class) +
  geom_point(aes(x = Urban, y = maxSumEAR)) +
  facet_wrap(. ~ Class, nrow = 3, scale = "free_y") +
  theme_bw() +
  theme(axis.text = element_blank())

ggplot(data = graph_class) +
  geom_point(aes(x = Agriculture, y = maxSumEAR)) +
  facet_wrap(. ~ Class, nrow = 3, scale = "free_y") +
  theme_bw() +
  theme(axis.text = element_blank())

```


