---
title: "Compare Mixture Methods"
output: 
  bookdown::word_document2:
    fig_caption: yes
    toc: no
editor_options: 
  chunk_output_type: console
---

# What chemicals do we lose?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.width = 7,
                      fig.height = 7)

library(toxEval)
library(tidyverse)
library(flextable)
source("R/mixtures/mix_script.R")

drake::loadd(chemicalSummary)

ACC_all = get_ACC(drake::readd(chem_info)$CAS) %>%
  remove_flags()

end_point_info <- end_point_info
cleaned_ep = clean_endPoint_info(end_point_info)

filtered_ep = filter_groups(cleaned_ep, 
                    groupCol = 'intended_target_family',
                    assays = c('ATG','NVS','OT',
                       'TOX21','CEETOX','APR',
                       'CLD','TANGUAY','NHEERL_PADILLA',
                       'NCCT_SIMMONS','ACEA'),
                    remove_groups = c('Background Measurement',
                                      'Undefined'))


```

## ToxCast Endpoint

```{r setupConcs, fig.height = 8, fig.cap="Concentration of chemicals with no ToxCast endpoints."}

ep_summary <- chemicalSummary %>% 
  group_by(chnm, CAS) %>% 
  summarise(n_endPoints = length(unique(endPoint)),
      n_sites_det = length(unique(shortName[EAR > 0])))

# What we lose:
no_ep_chem_info <- drake::readd(chem_info) %>% 
  filter(!(CAS %in% unique(ep_summary$CAS)))

no_ep_chem_data = drake::readd(all_data_fixed_cas) %>% 
  filter(CAS %in% no_ep_chem_info$CAS)


tox_list_no_ep <- list(chem_data = no_ep_chem_data,
                       chem_info = no_ep_chem_info,
                       chem_site = drake::readd(site_info),
                       benchmarks = drake::readd(benchmarks)) %>% 
  as.toxEval()

chemicalSummary_no_ep <- get_chemical_summary(tox_list = tox_list_no_ep)

plot_tox_boxplots(chemicalSummary_no_ep, category = "Chemical", x_label = "Concentration [ug/L]") 

```


## AOP

```{r aopData, fig.height=3, fig.cap="EARs of chemicals that we lose with current list of AOPs"}

AOP_crosswalk <- data.table::fread(here::here("data/supplemental/AOP_crosswalk_Dec_2018.csv")) %>%
  data.frame() %>%
  select(endPoint=Component.Endpoint.Name, ID=AOP.., KE = KE., everything()) %>%
  distinct()

aop_summary <- chemicalSummary %>% 
  left_join(select(AOP_crosswalk, endPoint, ID), by="endPoint") %>% 
  filter(!is.na(ID)) %>% 
  group_by(chnm, CAS) %>%
  summarise(n_AOP = length(unique(ID)),
            n_sites_det = length(unique(shortName[EAR > 0])))  

missing_chem <- ep_summary$chnm[!(ep_summary$chnm %in% unique(aop_summary$chnm))]
missing_cas <- ep_summary$CAS[!(ep_summary$CAS %in% unique(aop_summary$CAS))]

# What we lose:
no_aop_chem_info <- drake::readd(chem_info) %>% 
  filter(CAS %in% missing_cas)

no_aop_chem_data = drake::readd(all_data_fixed_cas) %>% 
  filter(CAS %in% no_aop_chem_info$CAS)

tox_list_no_AOPs <- list(chem_data = no_aop_chem_data,
                         chem_info = no_aop_chem_info,
                         chem_site = drake::readd(site_info),
                         exclude = drake::readd(exclude)) %>% 
  as.toxEval()


chemicalSummary_no_aop <- get_chemical_summary(
    tox_list = tox_list_no_AOPs,
    ACC = ACC_all, 
    filtered_ep = filtered_ep)

plot_tox_boxplots(chemicalSummary_no_aop, 
                  category = "Chemical")

```


## Gene targets

All measured chemicals with EARs have an associated gene target.

```{r gene}
gene_info <- select(end_point_info,
                    endPoint = assay_component_endpoint_name,
                    geneID = intended_target_gene_id,
                    geneName = intended_target_gene_name,
                    geneSymbol = intended_target_gene_symbol)

multiple_genes <- unique(gene_info$geneSymbol[grep(pattern = "\\|",
                                                   x = gene_info$geneSymbol)])

suppressWarnings({
gene_info_wide <- gene_info %>%
  mutate(orig_symbol = geneSymbol) %>% 
  separate(geneSymbol, into = c("a","b","c","d"),
           sep = "\\|") %>% 
  separate(geneID, into = c("IDa","IDb","IDc","IDd"),
           sep = "\\|") %>% 
  separate(geneName, into = c("Namea","Nameb","Namec","Named"),
           sep = "\\|")
})

gene_info_long <- gene_info_wide %>% 
  select(endPoint, geneID = IDa, geneSymbol = a, geneName = Namea) %>% 
  rbind(gene_info_wide %>% 
          select(endPoint, geneID = IDb, geneSymbol = b, geneName = Nameb) %>% 
          filter(!is.na(geneID))) %>% 
  rbind(gene_info_wide %>% 
          select(endPoint, geneID = IDc, geneSymbol = c, geneName = Namec) %>% 
          filter(!is.na(geneID))) %>% 
  rbind(gene_info_wide %>% 
          select(endPoint, geneID = IDd, geneSymbol = d, geneName = Named) %>% 
          filter(!is.na(geneID)))

gene <- select(gene_info_long,
               endPoint,
               gene = geneSymbol)

gene_summary <- chemicalSummary %>% 
    left_join(gene, by="endPoint") %>%  
    mutate(chnm = as.character(chnm)) %>% 
    filter(!is.na(gene)) %>% 
    group_by(chnm, CAS) %>%
    summarise(n_genes = length(unique(gene)),
              n_sites_det = length(unique(shortName[EAR > 0])))  

missing_chem_gene <- ep_summary$chnm[!(ep_summary$chnm %in% unique(gene_summary$chnm))]
missing_cas_gene <- ep_summary$CAS[!(ep_summary$CAS %in% unique(gene_summary$CAS))]

no_gene_chem_info <- drake::readd(chem_info) %>% 
    filter(CAS %in% missing_cas_gene)

no_gene_chem_data = drake::readd(all_data_fixed_cas) %>% 
    filter(CAS %in% no_gene_chem_info$CAS)

tox_list_no_path <- list(chem_data = no_gene_chem_data,
                         chem_info = no_gene_chem_info,
                         chem_site = drake::readd(site_info),
                         exclude = drake::readd(exclude)) %>% 
    as.toxEval()

# 
# chemicalSummary_no_gene <- get_chemical_summary(
#                    tox_list = tox_list_no_path,
#                    ACC = ACC_all, 
#                    filtered_ep = filtered_ep)
# 
# plot_tox_boxplots(chemicalSummary_no_gene,
#                   category = "Chemical")


```

## Panther

```{r pantherData, fig.cap="Chemicals we lose with panther database", fig.height=4.5}

panther <- data.table::fread("panther_data/joined_genes.csv",
                             data.table = FALSE) %>% 
  select(gene = gene_abbr,
         pathway_accession,
         pathway_name) %>% 
  left_join(gene,  by="gene")

panther_summary <- chemicalSummary %>% 
  left_join(panther, by="endPoint") %>%  
  filter(!is.na(gene)) %>% 
  filter(pathway_accession != "") %>% 
  group_by(chnm, CAS) %>%
  summarise(n_pathways = length(unique(pathway_accession)),
            n_sites_det = length(unique(shortName[EAR > 0])))  

missing_chem <- ep_summary$chnm[!(ep_summary$chnm %in% unique(panther_summary$chnm))]

missing_cas <- ep_summary$CAS[!(ep_summary$CAS %in% unique(panther_summary$CAS))]

# What we lose:
no_path_chem_info <- drake::readd(chem_info) %>% 
  filter(CAS %in% missing_cas)

no_path_chem_data = drake::readd(all_data_fixed_cas) %>% 
  filter(CAS %in% no_path_chem_info$CAS)

tox_list_no_path <- list(chem_data = no_path_chem_data,
                         chem_info = no_path_chem_info,
                         chem_site = drake::readd(site_info),
                         exclude = drake::readd(exclude)) %>% 
  as.toxEval()


chemicalSummary_no_aop <- get_chemical_summary(tox_list = tox_list_no_path,
                                               ACC = ACC_all, 
                                               filtered_ep = filtered_ep)

plot_tox_boxplots(chemicalSummary_no_aop, 
                  category = "Chemical")

```


# How to the top mixtures change?

## ToxCast

```{r toxMixSetup}
ear_thresh <- 0.001
n_site_thresh <- 10

EARsum_endpoint <- sum_endpoints(chemicalSummary,
                                 ear_cutoff = ear_thresh)

contributing_chems <- calc_contr_chems(EARsum_endpoint)

top_mixes <- top_mixes_fn(contributing_chems, n_site_thresh)

```

Using ear_thresh = `r ear_thresh`, and n_site_thresh = `r n_site_thresh`:

```{r toxMix}
top_flex <- flextable(select(top_mixes, 
                             contr_chems_st, endPoint,
                             n_samples, unique_sites)) 
top_flex <- autofit(top_flex)
# top_flex <- width(top_flex, j = 1, width = 4)
top_flex
```


## AOP

```{r aopMixSetup}

ear_thresh_aop <- 0.001
n_site_thresh_aop <- 10

aop_summary <- chemicalSummary %>% 
  left_join(select(AOP_crosswalk, 
                   endPoint, ID), by="endPoint") %>% 
  filter(!is.na(ID)) %>% 
  select(-endPoint) %>% 
  mutate(endPoint = as.character(ID))

EARsum_endpoint_aop <- sum_endpoints(aop_summary,
                                 ear_cutoff = ear_thresh_aop)

contributing_chems_aop <- calc_contr_chems(EARsum_endpoint_aop)

top_mixes_aop <- top_mixes_fn(contributing_chems_aop,
                              n_site_thresh_aop)

```

Using ear_thresh = `r ear_thresh_aop`, and n_site_thresh = `r n_site_thresh_aop`:

```{r toxMixAOP}
top_flex_aop <- flextable(select(top_mixes_aop, 
                             contr_chems_st, AOP_id = endPoint,
                             n_samples, unique_sites)) 

top_flex_aop <- autofit(top_flex_aop)
top_flex_aop <- width(top_flex_aop, j = 1, width = 4)
top_flex_aop
```

## Gene targets

```{r geneMixSetup}

ear_thresh_gene <- 0.001
n_site_thresh_gene <- 10

gene_summary <- chemicalSummary %>% 
  left_join(gene, by="endPoint") %>% 
  filter(!is.na(gene)) %>% 
  select(-endPoint) %>% 
  mutate(endPoint = as.character(gene))

EARsum_endpoint_gene <- sum_endpoints(gene_summary,
                                 ear_cutoff = ear_thresh_gene)

contributing_chems_gene <- calc_contr_chems(EARsum_endpoint_gene)

top_mixes_gene <- top_mixes_fn(contributing_chems_gene,
                              n_site_thresh_gene)

```

Using ear_thresh = `r ear_thresh_gene`, and n_site_thresh = `r n_site_thresh_gene`:

```{r toxMixGene}
top_flex_gene <- flextable(select(top_mixes_gene, 
                             contr_chems_st, Gene = endPoint,
                             n_samples, unique_sites)) 

top_flex_gene <- autofit(top_flex_gene)
top_flex_gene <- width(top_flex_gene, j = 1, width = 4)
top_flex_gene
```


## Panther

```{r pantherMixSetup}

ear_thresh_panther <- 0.0001
n_site_thresh_panther <- 10

panther_summary <- chemicalSummary %>% 
  left_join(panther, by="endPoint") %>%
  filter(!is.na(gene)) %>% 
  filter(pathway_accession != "") %>% 
  select(-endPoint) %>% 
  mutate(endPoint = pathway_name)

EARsum_endpoint_panther <- sum_endpoints(panther_summary,
                           ear_cutoff = ear_thresh_panther)

contributing_chems_panther <- calc_contr_chems(EARsum_endpoint_panther)

top_mixes_panther <- top_mixes_fn(contributing_chems_panther,
                              n_site_thresh_panther)

```

Using ear_thresh = `r ear_thresh_panther`, and n_site_thresh = `r n_site_thresh_panther`:

```{r toxMixPanther}
top_flex_panther <- flextable(select(top_mixes_panther, 
                             contr_chems_st, Panther_Name = endPoint,
                             n_samples, unique_sites)) 

top_flex_panther <- autofit(top_flex_panther)
top_flex_panther <- width(top_flex_panther, j = 1, width = 2)
top_flex_panther <- width(top_flex_panther, j = 2, width = 3)
top_flex_panther
```
