---
output: 
  pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.width = 10,
                      fig.height = 9)

library(tidyverse)
library(toxEval)
library(ggpubr)
library(cowplot)

source(file = "read_chemicalSummary.R")
source(file = "R/analyze/get_sites_ready.R")
source(file = "R/report/stack_plots.R")
source(file = "R/report/combo_plot2.R")
source(file = "R/analyze/explore_detection_limits.R")
source(file = "SI_site_mix_figure.R")

tox_list <- create_toxEval(file.path(Sys.getenv("PASSIVE_PATH"),                        "data","data_for_git_repo","clean",
                "passive.xlsx"))

site_info <- prep_site_list(tox_list$chem_site)
cas_final <- readRDS(file.path(Sys.getenv("PASSIVE_PATH"),
                               "data", "data_for_git_repo",
                               "clean",
                               "cas_df.rds"))


graphData_tox_det <- readRDS(file.path(Sys.getenv("PASSIVE_PATH"),
                               "data", "data_for_git_repo",
                               "clean","graphData_tox_det.rds"))


color_map <- class_colors(tox_list)
font_size <- 8

ordered_class <- levels(graphData_tox_det$Class)

levels(chemicalSummary$Class)[levels(chemicalSummary$Class) == "Food Additive/Plasticizer"] <- "Food Additive"
levels(chemicalSummary$Class)[levels(chemicalSummary$Class) == "Antimicrobial disinfectant"] <- "Antimicrobial"

ordered_class[ordered_class == "Food Additive/Plasticizer"] <- "Food Additive"
ordered_class[ordered_class == "Antimicrobial disinfectant"] <- "Antimicrobial"

tox_list$chem_info$Class[tox_list$chem_info$Class == "Food Additive/Plasticizer"] <- "Food Additive"
tox_list$chem_info$Class[tox_list$chem_info$Class == "Antimicrobial disinfectant"] <- "Antimicrobial"

name_key <- chemicalSummary %>% 
  select(CAS, chnm) %>% 
  distinct() %>% 
  left_join(select(cas_final, CAS, chnm_new = chnm), by = "CAS") %>% 
  arrange(chnm)

name_key$chnm_new[name_key$chnm_new == "Tris(1,3-dichloro-2-propyl)phosphate (TDCPP)"] <- "Tris(1,3-dichloro-2-propyl)phosphate"
name_key$chnm_new[name_key$chnm_new == "Tris(2-butoxyethyl)phosphate (TBEP)"] <- "Tris(2-butoxyethyl)phosphate"
name_key$chnm_new[name_key$chnm_new == "Tris(1-chloro-2-propyl)phosphate (TCPP)"] <- "Tris(1-chloro-2-propyl)phosphate"
name_key$chnm_new[name_key$chnm_new == "Tris(2-ethylhexyl)phosphate (TEHP)"] <- "Tris(2-ethylhexyl)phosphate"
                    
                    
levels(chemicalSummary$chnm) <- name_key$chnm_new[which(name_key$chnm == levels(chemicalSummary$chnm))]

classes <- unique(tox_list$chem_info$Class)

num_chem_to_keep <- 5

class_plots <- list()

site_info <- site_info %>% 
  arrange(`Short Name`)

levels(site_info$`Short Name`) <- paste0(levels(site_info$`Short Name`),
                                         " (",site_info$map_nm,")")

color_6 <- colorspace::rainbow_hcl(6, 
                                   start = -360, end = -55, 
                                   c = 100, l = 64)

for(i in ordered_class){
  
  chem_i <- chemicalSummary %>%
    dplyr::filter(Class == i,
                  EAR > 0)
  
  if(nrow(chem_i) == 0){
    next
  }
  
  orig_levels <- levels(chem_i$chnm)
  chem_i$chnm <- droplevels(chem_i$chnm)

  graphData <- graph_chem_data(chem_i)
  
  orderChem_df <- graphData %>%
    mutate(logEAR = log(meanEAR),
           chnm = as.character(chnm)) %>% 
    group_by(chnm) %>%
    summarise(max = max(logEAR[logEAR != 0], na.rm = TRUE)) %>%
    ungroup()
  
  orderChem_df$max[is.na(orderChem_df$max)] <- min(orderChem_df$max, na.rm = TRUE) - 1
  orderChem_df <- arrange(orderChem_df, max)

  new_levels <- orderChem_df$chnm %>% 
    rev()
  
  chem_i$chnm <- factor(as.character(chem_i$chnm), levels = new_levels)
  
  if(length(levels(chem_i$chnm)) > num_chem_to_keep+1){
    just_right <- levels(chem_i$chnm)[1:num_chem_to_keep]
    too_small <- levels(chem_i$chnm)[(num_chem_to_keep+1):length(levels(chem_i$chnm))]
    too_small_text <- paste0("Other [",length(too_small),"]")
    chem_i$chnm <- as.character(chem_i$chnm)
    chem_i$chnm[chem_i$chnm %in% too_small] <- too_small_text
    chem_i$chnm <- factor(chem_i$chnm, levels = c(just_right, too_small_text))
    
  }  

  color_i <- color_6[1:length(levels(chem_i$chnm))]
  names(color_i) <- levels(chem_i$chnm)
  
  if(grepl("Other",names(color_i)[length(color_i)])){
    color_i[length(color_i)] <- "grey77"
  }
  
  class_plots[[i]] <- whole_stack(chem_i, site_info, tox_list,
                            color_i, font_size, 
                            category = "Chemical",
                            title = i)
}

rel_widths <- c(3.25,3,rep(c(1,3),2), 1, 4)
general_lp <- c(0.05,0.87)

```


```{r fig1, fig.height=13}
create_DL_plot(tox_list)

```
Figure SI 1: Exposure activity ratios (EARs) at the detection level for chemicals monitored with passive samplers in Great Lakes tributaries, 2010-2014.

```{r fig2A}

plot_grid(
  class_plots[[ordered_class[1]]]$chem_count,
  class_plots[[ordered_class[1]]]$no_axis +
    theme(legend.position = general_lp,
          strip.text.y =  element_blank(),
          axis.title.x = element_blank()),
  class_plots[[ordered_class[2]]]$chem_count +
    theme(axis.text.y = element_blank(),
          axis.title.x = element_blank()),
  class_plots[[ordered_class[2]]]$no_axis +
    theme(legend.position = general_lp ,
          strip.text.y =  element_blank(),
          axis.title.x = element_blank()),
  class_plots[[ordered_class[3]]]$chem_count +
    theme(axis.text.y = element_blank(),
          axis.title.x = element_blank()),
  class_plots[[ordered_class[3]]]$no_axis +
    theme(legend.position = general_lp,
          strip.text.y =  element_blank(),
          axis.title.x = element_blank()),
  class_plots[[ordered_class[4]]]$chem_count +
    theme(axis.text.y = element_blank(),
          axis.title.x = element_blank()),
  class_plots[[ordered_class[4]]]$no_axis +
    theme(legend.position = c(0.015, 0.84),
          strip.text.y = 
            element_text(size = 1.25*font_size),
          axis.title.x = element_blank()),
  rel_widths = rel_widths,
  nrow = 1, labels = c("A")
)

```
\begin{center}
Sum of Maximum EAR\textsubscript{chem} Values
\end{center}

Figure SI 2-A: Maximum exposure-activity ratios (EAR) for `r paste0(tolower(ordered_class[1:4]), collapse = ", ")` computed from passive sampler chemistry data from 69 Great Lakes tributaries collected in 2010 and 2014 for chemicals included in the ToxCast database. [Site names are followed parenthetically by the map names from Figure 1; # Chems, number of chemicals with computed EAR values; EAR~chem~, exposure activity ratio]")

```{r fig2B}
plot_grid(
  class_plots[[ordered_class[5]]]$chem_count ,
  class_plots[[ordered_class[5]]]$no_axis +
    theme(legend.position = general_lp,
          strip.text.y =  element_blank(),
          axis.title.x = element_blank()),
  class_plots[[ordered_class[6]]]$chem_count +
    theme(axis.text.y = element_blank(),
          axis.title.x = element_blank()),
  class_plots[[ordered_class[6]]]$no_axis +
    theme(legend.position = general_lp,
          strip.text.y =  element_blank(),
          axis.title.x = element_blank()),
  class_plots[[ordered_class[7]]]$chem_count +
    theme(axis.text.y = element_blank(),
          axis.title.x = element_blank()),
  class_plots[[ordered_class[7]]]$no_axis +
    theme(legend.position = general_lp ,
          strip.text.y =  element_blank(),
          axis.title.x = element_blank()),
  class_plots[[ordered_class[8]]]$chem_count +
    theme(axis.text.y = element_blank(),
          axis.title.x = element_blank()),
  class_plots[[ordered_class[8]]]$no_axis +
    theme(legend.position = general_lp,
          strip.text.y = element_text(size = 1.25*font_size),
          axis.title.x = element_blank()),
  rel_widths = rel_widths,
  nrow = 1, labels = c("B")
)

```
\begin{center}
Sum of Maximum EAR\textsubscript{chem} Values
\end{center}

Figure SI 2-B: Maximum exposure-activity ratios (EAR) for `r paste0(tolower(ordered_class[5:8]), collapse = ", ")` computed from passive sampler chemistry data from 69 Great Lakes tributaries collected in 2010 and 2014 for chemicals included in the ToxCast database. [Site names are followed parenthetically by the map names from Figure 1; # Chems, number of chemicals with computed EAR values; EAR~chem~, exposure activity ratio]")

```{r fig2C}
plot_grid(
  class_plots[[ordered_class[9]]]$chem_count ,
  class_plots[[ordered_class[9]]]$no_axis +
    theme(legend.position = general_lp,
          strip.text.y =  element_blank(),
          axis.title.x = element_blank()),
  class_plots[[ordered_class[10]]]$chem_count +
    theme(axis.text.y = element_blank(),
          axis.title.x = element_blank()),
  class_plots[[ordered_class[10]]]$no_axis +
    theme(legend.position = general_lp,
          strip.text.y =  element_blank(),
          axis.title.x = element_blank()),
  class_plots[[ordered_class[11]]]$chem_count +
    theme(axis.text.y = element_blank()),
  class_plots[[ordered_class[11]]]$no_axis +
    theme(legend.position = c(0.025, 0.92) ,
          strip.text.y =  element_blank(),
          axis.title.x = element_blank()),
  class_plots[[ordered_class[12]]]$chem_count +
    theme(axis.text.y = element_blank(),
          axis.title.x = element_blank()),
  class_plots[[ordered_class[12]]]$no_axis +
    theme(legend.position = general_lp,
          strip.text.y = element_text(size = 1.25*font_size),
          axis.title.x = element_blank()),
  rel_widths = rel_widths,
  nrow = 1, labels = c("C")
)

```
\begin{center}
Sum of Maximum EAR\textsubscript{chem} Values
\end{center}

Figure SI 2-C: Maximum exposure-activity ratios (EAR) for `r paste0(tolower(ordered_class[9:12]), collapse = ", ")` computed from passive sampler chemistry data from 69 Great Lakes tributaries collected in 2010 and 2014 for chemicals included in the ToxCast database. [Site names are followed parenthetically by the map names from Figure 1; # Chems, number of chemicals with computed EAR values; EAR~chem~, exposure activity ratio]")

```{r fig2D}
plot_grid(
  class_plots[[ordered_class[13]]]$chem_count ,
  class_plots[[ordered_class[13]]]$no_axis +
    theme(legend.position = general_lp,
          strip.text.y = element_blank(),
          axis.title.x = element_blank()),
  class_plots[[ordered_class[14]]]$chem_count +
    theme(axis.text.y = element_blank(),
          axis.title.x = element_blank()),
  class_plots[[ordered_class[14]]]$no_axis +
    theme(legend.position = general_lp,
          strip.text.y = element_blank(),
          axis.title.x = element_blank()),
  class_plots[[ordered_class[15]]]$chem_count +
    theme(axis.text.y = element_blank(),
          axis.title.x = element_blank()),
  class_plots[[ordered_class[15]]]$no_axis +
    theme(legend.position = c(0.05, 0.92) ,
          strip.text.y =  element_blank(),
          axis.title.x = element_blank()),
  class_plots[[ordered_class[16]]]$chem_count +
    theme(axis.text.y = element_blank(),
          axis.title.x = element_blank()),
  class_plots[[ordered_class[16]]]$no_axis +
    theme(legend.position = c(0.05, 0.90),
          strip.text.y = element_text(size = 1.25*font_size),
          axis.title.x = element_blank()),
  rel_widths = rel_widths,
  nrow = 1, labels = c("D")
)

```
\begin{center}
Sum of Maximum EAR\textsubscript{chem} Values
\end{center}

Figure SI 2-D: Maximum exposure-activity ratios (EAR) for `r paste0(tolower(ordered_class[12:16]), collapse = ", ")` computed from passive sampler chemistry data from 69 Great Lakes tributaries collected in 2010 and 2014 for chemicals included in the ToxCast database. [Site names are followed parenthetically by the map names from Figure 1; # Chems, number of chemicals with computed EAR values; EAR~chem~, exposure activity ratio]")


```{r fig3}
plot_mixes()

```

Figure SI 3: Monitoring sites with exposure activity ratios > 10-3 at 10% or more of monitored sites for gene targets and associated chemical mixtures as defined in Table 1.
